Enabling GPU Virtualization in Cloud Environments

### 2015-10-26 paper.tex
background re-revisado

\section{\uppercase{Background}}
\label{sec:background}
\subsection{The rCUDA Framework}
\label{sec:rcuda}
{rCUDA}~\cite{toniparco} is a middleware that enables transparent access
to any NVIDIA GPU device present in a cluster from all compute
nodes. The GPUs can be accessed and shared between nodes, and a single node can use all the graphic accelerators
as if they were local.
These features focus in attaining higher accelerator utilization rates in the overall system while simultaneously reducing
resource, space, and energy requirements~\cite{energy14}.
rCUDA is structured following a client-server distributed
architecture: the client middleware runs in the same cluster node where the application demanding GPGPU
acceleration services is executed, providing a transparent replacement for the
native CUDA libraries. Furthermore, the server middleware is executed in the
cluster nodes from which the actual GPUs provide the requested GPGPU service.
To support a concurrent scenario where GPUs are shared between
processes\slash nodes, {rCUDA} manages separate device contexts for
each client application.

The {rCUDA} 5.0 client exposes the same interface as the regular NVIDIA
CUDA 6.5 release~\cite{cuda65}, including the runtime and driver
APIs as well as other commonly used libraries such as cuBLAS, cuFFT, cuSparse or cuRand.
Therefore, applications are not aware of the fact that they are being executed
on top of a virtualization layer.
To deal with new GPU programming models, {rCUDA} has been recently extended to accommodate 
directive-based models such as OmpSS~\cite{repara15} and OpenACC~\cite{cluster15}.

The integration of remote GPGPU virtualization with global
resource schedulers such as SLURM~\cite{sbacpad14} completes this powerful
technology, making accelerator-enabled clusters more flexible and
energy-efficient.

\subsection{Amazon Web Services}
\label{sec:aws}
AWS~\cite{aws} is a public cloud computing provider, 
composed of several services, such as 
 cloud-based computation, storage and other functionality, 
that enables organizations and/or individuals to deploy
services and applications on an on-demand basis. 
These services replace local IT infrastructure for companies and provide agility and instant elasticity matching 
perfectly with their software requirements.

From the point of view of HPC, AWS offers high performance facilities via 
instances equipped with GPUs and high performance network interconnection, which 
are common resources leveraged in this research field.

@@

OpenStack~\cite{OpenStack} is a cloud operating system that provides Infrastructure as a Service (IaaS). 
It is able to control large pools of compute, storage, and networking resources throughout a 
datacenter. All these resources are managed through a dashboard or an API that gives administrators 
control while empowering their users to provision resources through a web interface
or a command-line interface.  
OpenStack supports most available hypervisors and handles provisioning 
and life-cycle management of VMs.

The OpenStack architecture offers flexibility to create a custom cloud, with no proprietary hardware
or software requirements and the ability to integrate with legacy systems and third party technologies.

From the HPC perspective, OpenStack offers high performance configurations using
multiple supported hypervisors and different hardware architectures. Despite in OpenStack is possible to 
specify some extra parameters requesting special resources, like GPUs, choosing a host based on the existence
of a GPU is currently unsupported~\cite{OpenStackGPU}.

>>>>

\section{\uppercase{Background}}
\label{sec:background}
\subsection{The rCUDA Framework}
\label{sec:rcuda}
{rCUDA}~\cite{toniparco} is a middleware that enables transparent access
to any NVIDIA GPU device present in a cluster from all compute
nodes. The GPUs can also be shared among nodes, and a single node can use all the graphic accelerators
as if they were local.
These features aim to attain higher accelerator utilization rates in the overall system while simultaneously reducing
resource, space, and energy requirements~\cite{energy14}.
rCUDA is structured following a client-server distributed
architecture: the client middleware runs in the same cluster node where the application demanding GPGPU
acceleration services is executed, providing a transparent replacement for the
native CUDA libraries. Furthermore, the server middleware is executed in the
cluster nodes from which the actual GPUs provide the requested GPGPU services.
To support a concurrent scenario where GPUs are shared between
processes\slash nodes, {rCUDA} manages separate device contexts for
each client application.

The {rCUDA} 5.0 client exposes the same interface as the regular NVIDIA
CUDA 6.5 release~\cite{cuda65}, including the runtime and driver
APIs (Application Programming Interface) as well as other frequently used libraries such as cuBLAS, cuFFT, cuSparse or cuRand.
With this middleware, applications are not aware that they are executed
on top of a virtualization layer.
To deal with new GPU programming models, {rCUDA} has been recently extended to accommodate 
directive-based models such as OmpSs~\cite{repara15} and OpenACC~\cite{cluster15}.

The integration of remote GPGPU virtualization with global
resource schedulers such as SLURM~\cite{sbacpad14} completes this appealing 
technology, making accelerator-enabled clusters more flexible and
energy-efficient.

\subsection{Amazon Web Services}
\label{sec:aws}
AWS~\cite{aws} is a public cloud computing provider, 
composed of several services, such as 
 cloud-based computation, storage and other functionality, 
that enables organizations and/or individuals to deploy
services and applications on demand. 
These services replace company-owned local IT infrastructure and provide agility and instant elasticity matching 
perfectly with enterprise software requirements.

From the point of view of HPC, AWS offers high performance facilities via 
instances equipped with GPUs and high performance network interconnection.

@@

OpenStack~\cite{OpenStack} is a cloud operating system (OS) that provides Infrastructure as a Service (IaaS). 
OpenStack controls large pools of compute, storage, and networking resources throughout a 
datacenter. All these resources are managed through a dashboard or an API that gives administrators 
control while empowering their users to provision resources through a web interface
or a command-line interface.  
OpenStack supports most recent hypervisors and handles provisioning 
and life-cycle management of VMs.

The OpenStack architecture offers flexibility to create a custom cloud, with no proprietary hardware
or software requirements, and the ability to integrate with legacy systems and third party technologies.

From the HPC perspective, OpenStack offers high performance configurations with different hardware architectures. Despite in OpenStack it is possible to 
request special resources, such as GPUs, choosing a host based on the existence
of a GPU is currently unsupported~\cite{OpenStackGPU}.
